SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-10 docs/questioning.tex
"executable source code" became "source code injection", also grouped table I

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fancyhdr}
%\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{color}
\usepackage{url}
%\usepackage{moreverb}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{mathptmx}
\usepackage{dingbat}
\usepackage{pifont}
\usepackage{acronym}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{supertabular}
\usepackage{listings}
\usepackage{threeparttable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{multirow}
\usepackage{subfigure}

@@

@@

\author{
\IEEEauthorblockN{Dimitris Mitropoulos,\IEEEauthorrefmark{1}
Panos Louridas,\IEEEauthorrefmark{2},
Michalis Polychronakis,\IEEEauthorrefmark{1}
and Angelos D. Keromytis\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}
Network Security Lab\\
Department of Computer Scinence\\
Columbia University\\
\{dimitro, mikepo, angelos\}@cs.columbia.edu\\
\IEEEauthorrefmark{2}
Software Engineering and Security Lab\\
Department of Management Science and Technology\\
Athens University of Economics and Business\\
louridas@aueb.gr
}}

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
executable source code-driven injection attacks dominate the top five
positions. Malicious users find new ways to bypass defense mechanisms
by using a variety of techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we initially show how the above attacks can be seen from
a {\it common perspective}. We achieve this by providing an
exploitation model that highlights the following fact: the majority of
the steps needed to perform application attacks based on executable
source code injection are common.\footnote{Note that we leave out
  attacks that are based on binary code injection (i.e. buffer
  overflows --- see Section~\ref{sec:attacks}).} This is validated by
the fact that there are protection mechanisms that can detect more
than one of these attacks. Then, we {\it categorize} a selection of
these protection mechanisms. The selection comprises protection
mechanisms that detect such attacks and are presented in publications
cited more than twenty times, and some of the latest mechanisms that
have been presented in top security conferences that have been cited
less than twenty times. We included the second group so that we would
not penalize recent research. Finally, we {\it analyze} each mechanism based
on the following dimensions:
\begin{itemize}
\item {\it Diagnostic Performance} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Computational Performance} We examine
  the overhead of the mechanisms
  that affect the experience of a web user who actually
  uses the protected application.
\item {\it Ease of use} We examine if the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security} We examine how resilient is the system to
  attackers that know how it works and wants to circumvent it.
\item {\it Detection Locus} Based on our exploitation model
  we examine the location where a mechanism detects an attack.
\item {\it Availability} We check if the mechanism and its
  testbed are available on-line.
available 
\end{itemize}

@@

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on executable source code injection under a common
  perspective.
\item We categorize and analyze proposed defenses using a set of
  criteria that are important for building protection mechanisms.
\item We find that researchers often test the diagnostic performance
  of their mechanisms poorly. 
\item We also find that researchers do not often make their mechanisms
  available to the community, thus no one can either use their
  mechanism, or replicate their experiments.
\item We discuss how our observations may be the main reasons why the
  mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria for new mechanisms
  and how our work can help researchers to design new countermeasures.
\end{enumerate}

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{attack-tree-uml.pdf}
\end{center}
\caption{\label{fig:taxonomy}A taxonomy of executable source
code injection attacks.}
\end{figure}

Code injection attacks can be divided into two categories. The first
involves binary code and the second executable source code. An
extensive survey on binary code injection attacks can be found
in~\cite{LC03}. Specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others) have been
presented in~\cite{PB04}. The countermeasures used to detect such
defects have already been surveyed~\cite{YJP12} (many of them are also
included in a book:~\cite{DKZ12} -- Section 13.8). In the present work
we do not include binary code injection, focusing instead on defenses
that protect applications against attacks based on the
executable source code injection category.

\subsection{Executable Source Code Injection Vulnerabilities}

Figure~\ref{fig:taxonomy} presents a taxonomy of executable source
code injection attacks. Such attacks may involve source code, either of a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
executable source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there is a
corresponding tuple that includes the step number for
every attack. For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} also illustrates in which step
the defenses detect corresponding attacks.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on executable source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes the step number for
every attack. Note that {\sc xss} attacks can also involve
{\it phishing}. Every {\it locus} indicates a step where mechanisms
detect attacks. For every locus we provide the corresponding
references:
locus {\sc x}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
locus {\sc y}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
locus {\sc z}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
locus $\Omega$~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on executable source code
injectiion. For each approach we provide the references
that present corresponding mechanisms:
{\bf 1}~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\bf 2}~\cite{BK04,JB07,GC09,APKLM10},
{\bf 3}~\cite{BWS05,SW06},
{\bf 4}~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\bf 5}~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\bf 6}~\cite{BV08,LV09,SMS13}.}
\end{figure*}

@@

There exist three basic etiological approaches used to protect web
applications from executable source code injection attacks.

@@

    \caption{Comparison summary of mechanisms developed to counter application attacks based on executable source code injection.}

@@

@@

>>>>

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fancyhdr}
%\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{color}
\usepackage{url}
%\usepackage{moreverb}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{mathptmx}
\usepackage{dingbat}
\usepackage{pifont}
\usepackage{acronym}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{supertabular}
\usepackage{listings}
\usepackage{threeparttable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{array}

@@

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

@@

\author{
\IEEEauthorblockN{Dimitris Mitropoulos,\IEEEauthorrefmark{1}
Panos Louridas,\IEEEauthorrefmark{2}
Michalis Polychronakis,\IEEEauthorrefmark{1}
and Angelos D. Keromytis\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}
Network Security Lab\\
Department of Computer Scinence\\
Columbia University\\
\{dimitro, mikepo, angelos\}@cs.columbia.edu\\
\IEEEauthorrefmark{2}
Software Engineering and Security Lab\\
Department of Management Science and Technology\\
Athens University of Economics and Business\\
louridas@aueb.gr
}}

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
source code-driven injection attacks dominate the top five
positions. Malicious users find new ways to bypass defense mechanisms
by using a variety of techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we initially show how the above attacks can be seen from
a {\it common perspective}. We achieve this by providing an
exploitation model that highlights the following fact: the majority of
the steps needed to perform application attacks based on
source code injection are common.\footnote{Note that we leave out
  attacks that are based on binary code injection (i.e. buffer
  overflows --- see Section~\ref{sec:attacks}).} This is validated by
the fact that there are protection mechanisms that can detect more
than one of these attacks. Then, we {\it categorize} a selection of
these protection mechanisms. The selection comprises protection
mechanisms that detect such attacks and are presented in publications
cited more than twenty times, and some of the latest mechanisms that
have been presented in top security conferences that have been cited
less than twenty times. We included the second group so that we would
not penalize recent research. Finally, we {\it analyze} each mechanism based
on the following dimensions:
\begin{itemize}
\item {\it Diagnostic Performance} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Computational Performance} We examine
  the overhead of the mechanisms
  that affect the experience of a web user who actually
  uses the protected application.
\item {\it Ease of use} We examine if the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security} We examine how resilient is the system to
  attackers that know how it works and wants to circumvent it.
\item {\it Detection Locus} Based on our exploitation model
  we examine the location where a mechanism detects an attack.
\item {\it Availability} We check if the mechanism and its
  testbed are available on-line.
available 
\end{itemize}

@@

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on source code injection under a common
  perspective.
\item We categorize and analyze proposed defenses using a set of
  criteria that are important for building protection mechanisms.
\item We find that researchers often test the diagnostic performance
  of their mechanisms poorly. 
\item We also find that researchers do not often make their mechanisms
  available to the community, thus no one can either use their
  mechanism, or replicate their experiments.
\item We discuss how our observations may be the main reasons why the
  mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria for new mechanisms
  and how our work can help researchers to design new countermeasures.
\end{enumerate}

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{attack-tree-uml.pdf}
\end{center}
\caption{\label{fig:taxonomy}A taxonomy of source
code injection attacks.}
\end{figure}

Code injection attacks can be divided into two categories. The first
involves binary code and the second source code. An
extensive survey on binary code injection attacks can be found
in~\cite{LC03}. Specific advances in exploiting such vulnerabilities
(i.e. ``heap smashing", ``arc injection" and others) have already
been presented~\cite{PB04} and the countermeasures used to detect such
defects have already been surveyed~\cite{YJP12} (many of them are also
included in a book:~\cite{DKZ12} -- Section 13.8). In the present work
we do not include binary code injection, focusing instead on defenses
that protect applications against attacks based on the
source code injection category.

\subsection{Source Code Injection Vulnerabilities}

Figure~\ref{fig:taxonomy} presents a taxonomy of source
code injection attacks. Such attacks may involve source code, either of a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there is a
corresponding tuple that includes the step number for
every attack. For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} also illustrates in which step
the defenses detect corresponding attacks.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes the step number for
every attack. Note that {\sc xss} attacks can also involve
{\it phishing}. Every {\it locus} indicates a step where mechanisms
detect attacks. For every locus we provide the corresponding
references:
locus {\sc x}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
locus {\sc y}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
locus {\sc z}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
locus $\Omega$~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

\begin{figure*} [ht]
\begin{center}
\leavevmode
\includegraphics[scale=0.65]{defenses.pdf}
\end{center}
\caption{\label{fig:defenses}The basic categories of countermeasures
against web application attacks, based on source code
injectiion. For each approach we provide the references
that present corresponding mechanisms:
{\bf 1}~\cite{NSS06,JKK06a,KKVJ06,KJKV09,TNH07,RDWDE07,YCIS07,OWVS08,PSC09,ML10,DDHPJ10,PS11,VDDPJ11,BV08,LV09},
{\bf 2}~\cite{BK04,JB07,GC09,APKLM10},
{\bf 3}~\cite{BWS05,SW06},
{\bf 4}~\cite{HCF05,PB05,XBS06,NLC07,VFJKKV07,PMP11,SLMS14},
{\bf 5}~\cite{LLW02,HO05,HO06,HO05b,VMV05,JEP08,WPLKK09,MS09,MKS09,MKLS11},
{\bf 6}~\cite{BV08,LV09,SMS13}.}
\end{figure*}

@@

There exist three basic etiological approaches used to protect web
applications from source code injection attacks.

@@

    \caption{Comparison summary of mechanisms developed to counter application attacks based on source code injection.}

@@

    \thickhline

@@

  \thickhline
  \thickhline
